\section{Very Short Overview}

\subsection{Privacy and the Media}

\begin{enumerate}
    \item \textbf{Torts and the First Amendment}: PLF p. 52.
    \item \textbf{Information gatheriing}: intrusion upon seclusion, paprazzi, 
    video voyeurism.
    \item \textbf{Disclosure of truthful information}: public disclosure of 
    private facts, newsworthiness, First Amendment limitations,
    \item \textbf{Dissination of false or misleading information}: defamation, 
    false light, infliction of emotional distress.
    \item \textbf{Appropriation of name or likeness}: appropriation, right of 
    publicity, noncommercial use, imitators.
\end{enumerate}

\subsection{Privacy and Law Enforcement}

\begin{enumerate}
    \item \textbf{Fourth Amendment}: wiretapping, bugging, REOP, third party 
    doctrine, open fields/curtilage, canine sniffs.
    \item \textbf{Surveillance}: \S\ 605, Title III, ECPA, CALEA.
    \item \textbf{Digial searches and seizures}: computer searches, 
    email/\emph{Warshak}, Internet metadata collection, keylogging.
    \item \textbf{National security and foreign intelligence}: balancing 
    privacy and speech, FISA, emergency exceptions, the wall.
\end{enumerate}

\subsection{Health Privacy}

\begin{enumerate}
    \item \textbf{Confidentiality of medical information}: HIPAA (privacy and 
    security rules), HITECH Act.
    \item \textbf{Constitutional protection of medical information}: 
    informational/decisional, constitutional torts, compelling state interest 
    in disclosure, \emph{Westinghouse} balancing factors.
    \item \textbf{Genetic information}: state regulations, future diary, cheek 
    swabs of arrestees.
\end{enumerate}

\subsection{Privacy and Government Records}

\begin{enumerate}
    \item \textbf{Public access to government records}: public court records, 
    pseudonymous litigation, FOIA/exemptions, arrest records, police reports, 
    Megan's Law.
    \item \textbf{Government records of personal information}: Fair 
    Information Practices, Privacy Act, DPPA.
\end{enumerate}

\subsection{Privacy of Financial and Commercial Data}

\begin{enumerate}
    \item \textbf{Financial services industry}: FCRA, ``consumer report,'' GLB 
    Act, identity theft.
    \item \textbf{Commercial entities}: selling data, stalking, FTC 
    safeguards, VPPA, COPPA, defining PII.
    \item \textbf{Data security}: breach notification laws, actual harm, FTC 
    enforcement.
\end{enumerate}

\subsection{International Privacy Law}

\begin{enumerate}
    \item \textbf{OECD guidelines}.
    \item \textbf{European privacy protection}: personal sovereignty vs. 
    dignity, ECHR 8, ``wide margin of appreciation,'' EU data protection 
    directive.
\end{enumerate}
